---
title: "Sierpinski3d. Figure 2"
author: |-
  J.Garriga; F.Bartumeus\
  ICREA Movement Ecology Laboratory (CEAB-CSIC)\
  jgarriga@ceab.csic.es
date: "BigMap_4.5.0, `r format(Sys.Date(), '%b %Y')`"
output: html_document
---

```{r, echo = FALSE}
library(knitr)
library(kableExtra)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, }
library(bigMap)
```

Run pt-SNE on Sierpinski-3d with perplexity 2030 (99% of the data set size) and a decreasing range of thread-ratio. We show:

  + loss of local structure accuracy with decreasing thread-ratio;
  + decrease in running time (increasing speed);
  + improvement of the final embedding by means of a final compression (EFC);
  + using different perplexities in EFC enhances the structure at the corresponding scale while preserving the captured global structure;


```{r, out.width = '33%', fig.show = 'hold'}
# source aux. stuff (graph plot function)
source('../graphs.R')
# load edge matrix
E <- as.matrix(read.csv('../sierpinski3d.edg', sep = '', header = F))
```

### Load data

```{r, }
D <- as.matrix(read.csv('../sierpinski3d.mtx', sep = '', header = F))
```

### Compute betas

```{r, eval = F, results = F}
g <- bdm.init(D, dSet.name = 'S3D', is.distance = T, ppx = 2030, threads = 4)
```

### Run ptSNE (perplexity = 2030, decreasing thread-ratio)

```{r, eval = F, results = F}
g1 <- bdm.ptsne(D, g, threads = 1, layers = 1)
g2 <- bdm.ptsne(D, g, threads = 3, layers = 2)
g3 <- bdm.ptsne(D, g, threads = 4, layers = 2)
g4 <- bdm.ptsne(D, g, threads = 5, layers = 2)
g5 <- bdm.ptsne(D, g, threads = 6, layers = 2)
g6 <- bdm.ptsne(D, g, threads = 8, layers = 2)
```

```{r, eval = F}
g.list <- list(g1, g2, g3, g4, g5, g6)
save(g.list, file = './s3d_2030.RData')
```

```{r, echo = F}
load('./s3d_2030.RData')
```

### Output

Note that the global structure is more or less preserved (we are using a very high perplexity) but the local structure gets worst as the thread-ratio is decreased.

```{r, out.width = '33%', fig.show = 'hold'}
nulL <- lapply(g.list, function(g) graph.plot(g, E))
```

### hl-Correlation

In terms of HL-correlation the loss in structure (on average) is minimal.

```{r, eval = F}
g.list <- lapply(g.list, function(g) bdm.hlCorr(D, g, threads = 4))
save(g.list, file = './s3d_2030.RData')
```

```{r, }
hlTable <- sapply(g.list, function(g) summary(g$hlC))
hlTable <- t(round(hlTable, 4))
threadRatio <- sapply(g.list, function(g) g$ptsne$layers /g$ptsne$threads)
rownames(hlTable) <- round(threadRatio, 2)
knitr::kable(hlTable, caption = 'hl-Correlation by thread-ratio') %>%
  kable_styling(full_width = F)
```

### k-ary neighborhood preservation

In terms of kNP, the global structure (linAUC) is notably preserved (with the exception of g5 with thread-ratio = 0.33, blue line) but the loss in local structre (logAUC) is clearly perceptible.

```{r, eval = F}
g.list <- lapply(g.list, function(g) bdm.knp(D, g, threads = 4))
save(g.list, file = './s3d_2030.RData')
```

```{r, }
bdm.knp.plot(g.list, ppxfrmt = 0)
```

### Running Times

Note the correlation between decreasing thread-ratio and decreasing running time.

```{r, }
rTimes <- sapply(g.list, function(g) c(g$ppx$t[3], g$t$epoch, g$t$ptsne[3], sum(c(g$ppx$t[3], g$t$ptsne[3]))))
rTimes <- round(rTimes, 2)
threadRatio <- sapply(g.list, function(g) g$ptsne$layers /g$ptsne$threads)
colnames(rTimes) <- round(threadRatio, 2)
rownames(rTimes) <- c('betas', 'epoch', 'ptSNE', 'total')
knitr::kable(rTimes, caption = 'Computation times (s) by thread-ratio') %>%
  kable_styling(full_width = F)
```

*Run on: Intel(R) Xeon(R) CPU E31225 @ 3.10GHz, 4 cores, 16GB RAM.*

### Embedding final compression (EFC)

Let's take g6 (perplexity 2030 and lowest thread-ratio=0.25) and apply EFC with different compression perplexities.

```{r, eval = F}
# take the embedding with lowest thread-ratio
g.2030c <- g.list[6:6]
# compress with ppx = 1845 (0.90)
g.2030c <- bdm.efc(D, g.2030c, ppx = 1845, iters = 50, threads = 4)
# compress with ppx = 1025 (0.50)
g.2030c <- bdm.efc(D, g.2030c, ppx = 1025, iters = 50, threads = 4)
# compress with ppx = 205 (0.10)
g.2030c <- bdm.efc(D, g.2030c, ppx = 205, iters = 50, threads = 4)
# compress with ppx = 102 (0.05)
g.2030c <- bdm.efc(D, g.2030c, ppx = 102, iters = 50, threads = 4)
# compress with ppx = 20 (0.01)
g.2030c <- bdm.efc(D, g.2030c, ppx = 20, iters = 50, threads = 4)
save(g.2030c, file = './s3d_2030c.RData')
```

```{r, echo = F}
load('./s3d_2030c.RData')
```

### EFC output vs. output from Fig.1 (with thread-ratio=1.0)

```{r, }
# load embeddings with thread-ratio = 1.0 (from Fig.1)
load('../pt-SNE/glist.RData')
g.threadRatio1 <- g.list[c(13, 11, 7, 3, 2, 1)]
```

```{r, out.width = '45%', fig.show = 'hold'}
g <- g.threadRatio1[[1]]
graph.plot(g, E, title = 'ppx=2030, thread-ratio=1.0')
g <- g.2030c[[1]]
graph.plot(g, E, title = 'ppx=2030, thread-ratio=0.25')
```

Note that decreasing EFC perplexities lead to the same scale resolutions we got using these same perplexities and thread-reatio = 1.0 (outputs shown in Fig.1), but the global structure we already captured is preserved.

```{r, out.width = '45%', fig.show = 'hold'}
nulL <- lapply(2:6, function(i) {
  g <- g.threadRatio1[[i]]
  graph.plot(g, E, title = paste('ppx=', g$ppx$ppx, ', thread-ratio=1.0', sep = ''))
  g <- g.2030c[[i]]
  graph.plot(g, E, title = paste('ppx=2030, thread-ratio=0.25, +efc.', g$ppx$ppx, sep = ''))
})
```

### EFC hl-Correlation

The HL-correlation shows a decrease as we enhanced local structure while preserving global structure because short/long distances are not equally preserved.

```{r, eval = F}
g.2030c <- lapply(g.2030c, function(g) bdm.hlCorr(D, g, zSampleSize = 1000, threads = 4))
```

```{r, }
hlTable <- sapply(g.2030c, function(g) summary(g$hlC))
hlTable <- t(round(hlTable, 4))
efc.ppx <- paste('efc.', sapply(g.2030c, function(g) g$ppx$ppx), sep = '')
rownames(hlTable) <- efc.ppx
knitr::kable(hlTable, caption = 'hl-Correlation by EFC-perplexity') %>%
  kable_styling(full_width = F)
```

### EFC k-ary neighborhood preservation

In terms of kNP we clearly observe that while preserving the global structure (linAUC), EFC results in a clear increase of local structure (logAUC).

```{r, eval = F}
g.2030c <- lapply(g.2030c, function(g) bdm.knp(D, g, k.max = NULL, sampling = 0.9, threads = 4))
```

```{r, }
bdm.knp.plot(g.2030c, ppxfrmt = 0)
```

### Running Times

The running time cost of EFC is obviously high, but a few iterations are usually enough to show some imprevement in the definition of the local structure.

```{r, echo = F}
# trick ;)
g.2030c[[1]]$t$efc <- c(0, 0, 0)
```

```{r, }
rTimes <- sapply(g.2030c, function(g) c(g$ppx$t[3], g$t$epoch, g$t$ptsne[3], g$t$efc[3], sum(c(g$ppx$t[3], g$t$ptsne[3], g$t$efc[3]))))
rTimes <- round(rTimes, 2)
colnames(rTimes) <- sapply(g.2030c, function(g) g$ppx$ppx)
rownames(rTimes) <- c('betas', 'epoch', 'ptSNE', 'EFC', 'total')
knitr::kable(rTimes, caption = 'Computation times (s) by EFC-perplexity') %>%
  kable_styling(full_width = F)
```

*Run on: Intel(R) Xeon(R) CPU E31225 @ 3.10GHz, 4 cores, 16GB RAM.*
