---
title: "Primes1M. pt-SNE"
author: |-
  J.Garriga; F.Bartumeus\
  ICREA Movement Ecology Laboratory (CEAB-CSIC)\
  jgarriga@ceab.csic.es
date: "bigMap_4.5.0 `r format(Sys.Date(), '%b %Y')`"
output: html_document
---

```{r, echo = FALSE}
library(knitr)
library(kableExtra)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, }
library(bigMap)
# load aux. stuff 
source('../primes.R')
```

### Load data

```{r, }
load('../P1M.RData')
```

### Run pt-SNE

### Load data

The Primes1M data set is a structured representation of the first 10^6 integers based on their prime factor decomposition. This is a highly sparse matrix with 78498 dimensions (primes lower than 10^6) where the values are the prime factorisation powers. We use a compact matrix format where odd columns hold the values of the columns (primes) in the original matrix and even columns hold the powers themselves. We also include number 0 in the first row with all zeros, but we do not incude number 1. So rows 2, 3, 4, ... correspond to numbers 2, 3, 4, ... This results in a matrix with 10^6 rows and 14 columns.
  
```{r, }
load('../P1M.RData')
```

This data set is inspired by the work of John Williamson https://johnhw.github.io/umap_primes/index.md.html, though we note some differences:

  + in Williamson's work any factor is represented as either a zero or a one, just indicating divisibility by that factor, while in our approach we use the actual power for each factor to have unique representations for each integer;
  
  + in Williamson's work affinities are based on cosine similarity, while we use euclidean distances.

### Run pt-SNE

We use the script below to run pt-SNE on this data set using a HPC platform with 101 cores and 4GB/core,

```
# ./bm450/P1M_start.R: 
# Run pt-SNE on Primes1M
# Compute kNP and HL-correlation

# +++ load package
library(bigMap)

# +++ load data
load('./P1M.RData')

# +++ start MPI cluster
mpi.cl <- bdm.mpi.start(threads)
if (is.null(mpi.cl)) return()

# +++ range of perplexities
ppx.list <- c(5000, 10000, 25000, 50000)

# +++ run init (compute betas)
m <- bdm.init(P1M, dSet.name = 'P1M', is.sparse = T, ppx = ppx.list, threads = 101, mpi.cl = mpi.cl)
save(m, file = './P1M_init.RData')

# +++ run ptSNE
m.list <- bdm.ptsne(P1M, m, theta = 0.33, threads = 101, mpi.cl = mpi.cl, layers = 2)

# +++ kNP + hlC
m.list <- lapply(m.list, function(m)
{
  m <- bdm.knp(NULL, m, k.max = 500000, sampling = 0.25, threads = threads, mpi.cl = mpi.cl)
  m <- bdm.hlCorr(NULL, m, threads = threads, mpi.cl = mpi.cl)
  m
})

# +++ save 
save(m.list, file = './bm450/P1M_ptSNE.RData')

# +++ stop cluster
bdm.mpi.stop(mpi.cl)

```

Submit:

```
$ qsub -pe ompi 101 -l h_vmem=4G Rsnow ~/bm450/P1M_start.R
```
### Load output

```{r, }
# load m.list 
load('./P1M_ptsne.RData')
```

### Output

```{r, out.width = '50%', fig.show = 'hold'}
#nulL <- lapply(m.list, function(m) primes.plot(P1M, m))
```

### hl-Correlation

```{r, eval = F}
hlTable <- sapply(m.list, function(m) summary(m$hlC)[4])
hlTable <- matrix(round(hlTable, 4), nrow = 1)
colnames(hlTable) <- sapply(m.list, function(m) m$ppx$ppx)
rownames(hlTable) <- c('<hlC>')
knitr::kable(hlTable, caption = 'hl-Correlation') %>%
  kable_styling(full_width = F)
```

### k-ary neighborhood preservation

```{r, }
#bdm.knp.plot(m.list, ppxfrmt = 0)
```

### Running Times

```{r, }
rTimes <- sapply(m.list, function(m) c(m$ppx$t[[3]], m$t$ptsne[[3]]))
rTimes <- round(rbind(rTimes, apply(rTimes, 2, sum)) /60, 2)
colnames(rTimes) <- sapply(m.list, function(m) m$ppx$ppx)
rownames(rTimes) <- c('betas', 'pt-SNE', 'total')
knitr::kable(rTimes, caption = 'Computation times (min)') %>%
  kable_styling(full_width = F)
```

*Run on: Intel(R) Xeon(R) CPU E5-2650 v3 2.30GHz, 32Mb cache, 101 cores, 4GB/core RAM.*

